Comparing and combining network-guided biomarker discovery to discover genetic susceptibility mechanisms to breast cancer on the GENESIS study

### 2021-01-18 manuscript/plos/review_2.tex
replaced betweenness by normalized betweenness

% Title must be 250 characters or less.
\begin{flushleft}
{\Large
\textbf\newline{How can biological networks boost GWAS? A case study on familial breast cancer susceptibility} % Please use "sentence case" for title and headings (capitalize only the first word in a title (or heading), the first word in a subtitle (or subheading), and any proper nouns).
}
\newline
% Insert author names, affiliations and corresponding author email (do not include titles, positions, or degrees).
\\
Héctor Climente-González\textsuperscript{1,2,3,4*}, 
Christine Lonjou\textsuperscript{1,2,3}, 
Fabienne Lesueur\textsuperscript{1,2,3},
GENESIS study group\textsuperscript{\textpilcrow},
Dominique Stoppa-Lyonnet\textsuperscript{5,6,7\textcurrency}, 
Nadine Andrieu\textsuperscript{1,2,3}, 
Chloé-Agathe Azencott\textsuperscript{3,1,2}
\\
\bigskip
\textbf{1} Institut Curie, PSL Research University, F-75005 Paris, France;\\
\textbf{2} INSERM, U900, F-75005 Paris, France;\\
\textbf{3} MINES ParisTech, PSL Research University, CBIO-Centre for Computational Biology, F-75006 Paris, France;\\
\textbf{4} RIKEN Center for Advanced Intelligence Project (AIP), Tokyo, Japan; \\
\textbf{5} Service de Génétique, Institut Curie, F-75005 Paris, France;\\
\textbf{6} INSERM, U830, F-75005 Paris, France;\\
\textbf{7} Université Paris Descartes.\\
\bigskip

@@

\end{flushleft}
% Please keep the abstract below 300 words
\section*{Abstract}
Genome-wide association studies (GWAS) explore the genetic causes of complex diseases. However, they ignore the biological context of the SNPs and genes under study. To address this, one can seek susceptibility genes that are functionally related, using networks to integrate this functional information. Many such network methods exist, stemming from different mathematical frameworks, pre-processing steps, and assumptions about the susceptibility mechanism. Unsurprisingly, their results are also very different. To sort through this diversity, we selected six network methods and applied them to GENESIS, a GWAS on French women with non-BRCA familial breast cancer. First, we verified that network methods recovered more interesting and interpretable genes than standard GWAS. But, as expected, these results were notably different. To address this heterogeneity, we combined them into a \emph{consensus}. The key gene in the consensus solution was \emph{COPS5}, a gene related to multiple cancer hallmarks. Another drawback was that network methods were unstable, selecting very different genes on different subsamples of GENESIS. Therefore, we proposed a \emph{stable consensus} solution formed by the 68 genes most consistently selected in multiple subsamples. Importantly, this solution was also enriched in known breast cancer susceptibility genes (\emph{BLM}, \emph{CASP8}, \emph{CASP10}, \emph{DNAJC1}, \emph{FGFR2}, \emph{MRPS30}, and \emph{SLC4A7}, P\=/value~=~$3 \times 10^{-4}$). The most central gene was \emph{CUL3}, which regulates several genes linked to cancer progression. Lastly, we evaluated the biases of each method and the impact of the parameters on the outcome. In general, network methods preferred highly central genes, even in random rewirings that stripped networks of any biological meaning. In conclusion, we showed the advantages of network-guided GWAS, characterized their shortcomings, and provided strategies to address them. To compute the consensus networks, implementations of all six methods are available at \url{https://github.com/hclimente/gwas-tools}.

@@

In this work, we studied the relevance of network-based approaches to genetics by applying these six network methods to GWAS data. They use different interpretations of the omnigenic model and provide a representative view of the field. We worked on the GENESIS dataset \cite{sinilnikova_genesis:_2016}, a study on familial breast cancer conducted in the French population. After a classical GWAS approach, we used these network methods to identify additional breast cancer susceptibility genes. Lastly, we compared the solutions obtained by the different methods and computed their intersection to obtain consensus solutions of predisposition to familial breast cancer that addressed their shortcomings.

@@

\subsubsection{Mathematical notations}
\label{methods:notation}
In this article, we used undirected, vertex-weighted networks, or graphs, $G = (V,E,w)$. $V = \{v_{1}, \dots{}, v_{n}\}$ refers to the vertices, with weights $w: V \rightarrow \mathbb{R}$. Equivalently, $E \subseteq \{\{x,y\} | x,y \in V \wedge x \neq y\}$ refers to the edges. When referring to a subnetwork $S$, $V_{S}$ is the set of nodes in $S$ and $E_{S}$ is the set of edges in $S$. A special case of subgraphs are \emph{connected} subgraphs, which occur when every node in the subgraph can be reached from any other node.

Nodes can be described by properties provided by the topology of the graph. We focused on two of those: degree centrality and betweenness centrality. The degree centrality, or simply emph{degree}, is the number of edges that a node has. The betweenness centrality, or \emph{betweenness}, is the number of times a node participates in shortest paths between two other nodes.

We also used two matrices that describe two different properties of a graph. Both matrices are square and have as many rows and columns as nodes are in the network. The element $(i,j)$ hence represents a relationship between $v_i$ and $v_j$. The \emph{adjacency matrix} $W_G$ contains a 1 when the corresponding nodes are connected, and 0 otherwise; its diagonal is zero. The \emph{degree matrix} $D_G$ is a diagonal matrix that contains the degree of the different nodes.

@@

Rewiring the PPIN while preserving the number of edges of each gene allowed to study the impact of the topology on the output of network methods. Indeed, the edges lose their biological meaning while the topology of the network is conserved. We produced 100 such rewirings by randomly swapping edges in the PPIN. We still scored the genes as described in Section~\ref{methods:methods}. We applied only four methods on the rewirings: heinz, dmGWAS, LEAN, and SigMod. We excluded HotNet2 and SConES since they took notably longer to run. As on the real PPIN, LEAN did not produce significant results in the rewirings.

@@

The topologies of the five solutions differed as well (Fig~\ref{fig:solution_overview}C), as measured by the median centrality and the number of connected components (Table~\ref{tab:gene_solutions}). Three methods yielded more than one connected component: SConES, as described above, SigMod, and HotNet2. HotNet2 produced 135 subnetworks, 115 of which have fewer than five genes. The second largest subnetwork (13 nodes) contained the two breast cancer susceptibility genes \emph{CASP8} and \emph{BLM} (Section~\ref{methods:bcac}).

\begin{table}[!ht]
  \begin{adjustwidth}{-2.25in}{0in} % Comment out/remove adjustwidth environment if table fits in text column.
  \centering
  \caption{
  {\bf Summary statistics on the solutions of multiple network methods on the PPIN. The first row contains the summary statistics on the whole PPIN.}}
  \begin{tabular}{lrrrrrr}
  {\bf Network } & {\bf \# genes } & {\bf \# edges } & {\bf \# components } & {\bf \mean{Betweenness} } & {\bf \median{P}\textsubscript{gene} } & {\bf \# genes in consensus } \\
  \thickhline
  HINT HT           & 13\,619 & 142\,541  & 15  & 16\,706   & 0.46  & 93/93 \\
  \hline
  dmGWAS            & 194     & 450       & 1   & 49\,115   & 0.19  & 55/93 \\
  heinz             & 4       & 3         & 1   & 113\,633  & 0.001 & 4/93  \\
  HotNet2           & 440     & 374       & 130 & 7\,739    & 0.048 & 63/93 \\
  LEAN              & 0       & 0         & 0   & -         & -     & 0/93  \\
  SConES GI         & 0 (1)   & 0         & 0   & -         & -     & 0/93  \\
  SigMod            & 142     & 249       & 11  & 92\,603   & 0.008 & 84/93 \\
  \hline
  Consensus         & 93      & 186       & 21  & 50\,737   & 0.006 & 93/93 \\
  Stable consensus  & 68      & 49        & 32  & 94\,854   & 0.005 & 43/93 \\
  \end{tabular}
  \begin{flushleft} \textbf{\# genes}:~number of genes selected out of those that are part of the PPIN; for SConES GI, the total number of genes, including RNA genes, was added in parentheses. \textbf{\# components}:~number of connected components. \textbf{\mean{Betweenness}}:~median betweenness of the selected genes in the PPIN. \textbf{\median{P}\textsubscript{gene}}:~median VEGAS2 P\=/value of the selected genes. \textbf{\# genes in consensus}:~number of genes in common between the method's solution and the 93 genes in the consensus solution.
  \end{flushleft}
  \label{tab:gene_solutions}
  \end{adjustwidth}
\end{table}

@@

Despite their shared properties, the differences between the solutions suggested that each of them captured different aspects of cancer susceptibility. Indeed, out of the 668 genes that were selected by at least one method, only 93 were selected by at least two, 20 by three, and none by four or more. Encouragingly, the more methods selected a gene, the higher its association score to the phenotype (Fig~\ref{fig:solution_overview}B), a relationship that plateaued at 2. Hence, to leverage their strengths and compensate for their respective weaknesses, we built a consensus solution using the genes shared among at least two solutions (Section~\ref{methods:methods}). This solution (Fig~\ref{fig:consensus}) contained 93 genes and exhibited the aforementioned properties of the individual solutions: enrichment in breast cancer susceptibility genes and higher betweenness centrality than the rest of the genes.

@@

\begin{figure}[!ht]
  \centering
  \includegraphics[width=.8\linewidth]{./figures/figure_4.pdf}
  \caption{\textbf{Drawbacks encountered when using network methods.} \textbf{(A)}~DmGWAS solution, with the genes colored according to the -log$_{10}$ of their P\=/value. \textbf{(B)}~Number of times a gene was selected by either dmGWAS, heinz, LEAN, or SigMod in 100 rewirings of the PPIN (Section~\ref{methods:pathway_enrichment}) and its centrality degree. \textbf{(C)}~Centrality degree and -log\textsubscript{10} of the VEGAS2 P\=/value in BCAC for each of the nodes in the PPIN. We highlighted the genes selected by each method and the ones selected by more than one (``Consensus''). We labeled the three most central genes that were picked by any method. \textbf{(D)}~Overlap between the solutions of SConES GS, GM, or GI. Barplots are colored based on whether the SNPs map to a gene or not (Section~\ref{methods:snp2gene}).}
  \label{fig:issues}
  \end{figure}

Another peculiarity of network methods was their relationship to degree centrality. We studied random rewirings of the PPIN that preserved node centrality (Section~\ref{methods:rewiring}). In this setting, network methods favored central genes (Fig~\ref{fig:issues}B) even though highly central genes often had no association to breast cancer susceptibility (Fig~\ref{fig:issues}C). We found this bias especially in SigMod (\nameref{sfig:rewirings}), which selected three highly central, unassociated genes in both the PPIN and in many of the random rewirings: \emph{COPS5}, \emph{CUL3}, and \emph{FN1}. However, as we showed in Section~\ref{results:consensus} and will show in \ref{results:stable-consensus}, there is evidence in the literature of the contribution of the first two to breast cancer susceptibility. With regards to \emph{FN1}, it encodes a fibronectin, a protein of the extracellular matrix involved in cell adhesion and migration. Overexpression of \emph{FN1} has been observed in breast cancer \cite{Ioachim2002}, and it anticorrelates with poor prognosis in other cancer types \cite{Yi2016,Sponziello2016}.

@@

Most network methods, including the consensus, were highly unstable (Fig~\ref{fig:benchmark}B), raising questions about the results' reliability. We built a new, \emph{stable consensus} solution using the genes selected most often across the 30 solutions obtained by running the 6 methods on 5 different splits of the data (Section~\ref{methods:benchmark}). Such a network should capture the subnetworks more often found altered, and hence should be more resistant to noise. We used only genes selected in at least 7 solutions, which corresponded to 1\% of all genes selected at least once. The resulting stability-based consensus was composed of 68 genes (Fig~\ref{fig:stable-consensus}). This network shared most of the properties of the consensus: breast cancer susceptibility genes were overrepresented (P\=/value = 3 \texttimes{} 10\textsuperscript{-4}), as well as genes involved in mitochondrial translation and the attenuation phase (adjusted P\=/values 0.001 and 3 \texttimes{} 10\textsuperscript{-5} respectively); the selected genes were more central than average (P\=/value = 1.1 \texttimes{} 10\textsuperscript{-14}); and a considerable number of nodes (19) were isolated (Fig~\ref{fig:stable-consensus}B and C).

@@

\paragraph*{S5 Fig.}
\label{sfig:rewirings}
\textbf{Number of times a gene was selected by either dmGWAS, heinz, or SigMod in 100 rewirings of the PPIN (Section~\ref{methods:rewiring}) and its centrality degree.} This figure is equivalent to Fig~\ref{fig:issues}B, split by method. 
%\includegraphics[width=.9\linewidth]{./figures/sfigure_5.pdf

@@

We wish to thank Om Kulkarni for helpful discussion on gene-based GWAS and PPIN databases, and the genetic epidemiology platform (the PIGE, Plateforme d'Investigation en Génétique et Epidemiologie: S. Eon-Marchais, M. Marcou, D. Le Gal, L. Toulemonde, J. Beauvallet, N. Mebirouk, E. Cavaciuti), the biological resource center (S. Mazoyer, F. Damiola, L. Barjhoux, C. Verny-Pierre, V. Sornin). We wish to pay tribute to Olga M. Sinilnikova, one of the initiators and principal investigators of GENESIS, and who died prematurely on June 30, 2014.

>>>>

% Title must be 250 characters or less.
\begin{flushleft}
{\Large
\textbf\newline{How can biological networks boost GWAS? A case study on susceptibility to familial breast cancer} % Please use "sentence case" for title and headings (capitalize only the first word in a title (or heading), the first word in a subtitle (or subheading), and any proper nouns).
}
\newline
% Insert author names, affiliations and corresponding author email (do not include titles, positions, or degrees).
\\
Héctor Climente-González\textsuperscript{1,2,3,4*}, 
Christine Lonjou\textsuperscript{1,2,3}, 
Fabienne Lesueur\textsuperscript{1,2,3},
GENESIS study group\textsuperscript{\textpilcrow},
Dominique Stoppa-Lyonnet\textsuperscript{5,6,7\textcurrency}, 
Nadine Andrieu\textsuperscript{1,2,3}, 
Chloé-Agathe Azencott\textsuperscript{3,1,2}
\\
\bigskip
\textbf{1} Institut Curie, PSL Research University, F-75005 Paris, France;\\
\textbf{2} INSERM, U900, F-75005 Paris, France;\\
\textbf{3} MINES ParisTech, PSL Research University, CBIO-Centre for Computational Biology, F-75006 Paris, France;\\
\textbf{4} RIKEN Center for Advanced Intelligence Project (AIP), Tokyo, Japan; \\
\textbf{5} Service de Génétique, Institut Curie, F-75005 Paris, France;\\
\textbf{6} INSERM, U830, F-75005 Paris, France;\\
\textbf{7} Université Paris Descartes.\\
\bigskip

@@

\end{flushleft}
% Please keep the abstract below 300 words
\section*{Abstract}
Genome-wide association studies (GWAS) explore the genetic causes of complex diseases. However, they ignore the biological context of the SNPs and genes under study. To address this shortcoming, one can seek susceptibility genes that are functionally related, employing networks to integrate such functional information. Many such network methods exist, each stemming from different mathematical frameworks, pre-processing steps, and assumptions about the susceptibility mechanism. Unsurprisingly, their results are also disparate. To sort through this diversity, we selected six network methods and applied them to GENESIS, a GWAS on French women with non-BRCA familial breast cancer. First, we verified that network methods recovered more interesting and interpretable genes than a standard GWAS. Nevertheless, as expected, these results were notably different. We addressed this heterogeneity studying their overlap, computing what we called the \emph{consensus}. The key gene in this consensus solution was \emph{COPS5}, a gene related to multiple cancer hallmarks. Another drawback of network methods was their instability, selecting very different genes on different subsamples of GENESIS. Therefore, we proposed a \emph{stable consensus} solution formed by the 68 genes most consistently selected across multiple subsamples. Importantly, this solution was also enriched in genes associated with breast cancer susceptibility (\emph{BLM}, \emph{CASP8}, \emph{CASP10}, \emph{DNAJC1}, \emph{FGFR2}, \emph{MRPS30}, and \emph{SLC4A7}, P\=/value~=~$3 \times 10^{-4}$). The most connected gene was \emph{CUL3}, a regulator of several genes linked to cancer progression. Lastly, we evaluated the biases of each method and the impact of the parameters on the outcome. In general, network methods preferred highly connected genes, even after random rewirings that stripped the connections of any biological meaning. In conclusion, we presented the advantages of network-guided GWAS, characterized their shortcomings, and provided strategies to address them. To compute the consensus networks, implementations of all six methods are available at \url{https://github.com/hclimente/gwas-tools}.

@@

In this work, we studied the relevance of network-based approaches to genetics by applying these six network methods to GWAS data. They use different interpretations of the omnigenic model and provide a representative view of the field. We worked on the GENESIS dataset \cite{sinilnikova_genesis:_2016}, a study on familial breast cancer conducted in the French population. After a classical GWAS approach, we used these network methods to identify additional breast cancer susceptibility genes. Lastly, we compared the solutions obtained by the different methods and studied their intersection to obtain consensus solutions of predisposition to familial breast cancer that addressed their shortcomings.

@@

\subsubsection{Mathematical notations}
\label{methods:notation}
In this article, we used undirected, vertex-weighted networks, or graphs, $G = (V,E,w)$. $V = \{v_{1}, \dots{}, v_{n}\}$ refers to the vertices (or nodes), with weights $w: V \rightarrow \mathbb{R}$. Equivalently, $E \subseteq \{\{x,y\} | x,y \in V \wedge x \neq y\}$ refers to the edges. When referring to a subnetwork $S$, $V_{S}$ is the set of nodes in $S$ and $E_{S}$ is the set of edges in $S$. A special case of subgraphs are \emph{connected} subgraphs, which occur when every node in the subgraph can be reached from any other node.

Nodes can be described by properties provided by the topology of the graph. We focused on the betweenness centrality, or \emph{betweenness}: the number of times a node participates in a shortest path between two other nodes. We normalized the betweenness by dividing it by $\frac {(N-1)(N-2)} {2}$, where $N$ is the number of nodes.

We also used two matrices that describe two different properties of a graph. Both matrices are square and have as many rows and columns as nodes are in the network. The element $(i,j)$ hence represents a relationship between $v_i$ and $v_j$. The \emph{adjacency matrix} $W_G$ contains a 1 when the corresponding nodes are connected, and 0 otherwise; its diagonal is zero. The \emph{degree matrix} $D_G$ is a diagonal matrix that contains the number of edges of each node.

@@

Rewiring the PPIN while preserving the number of edges of each gene allowed to study the impact of the topology on the output of network methods. Indeed, the edges lose their biological meaning while the topology of the network is conserved. We produced 100 such rewirings by randomly swapping edges in the PPIN. We still scored the genes as described in Section~\ref{methods:methods}. We applied only four methods on the rewirings: heinz, dmGWAS, LEAN, and SigMod. We excluded HotNet2 and SConES since they took notably longer to run. As on the real PPIN, LEAN did not produce significant results on the rewirings either.

@@

The topologies of the five solutions differed as well (Fig~\ref{fig:solution_overview}C), as measured by the median betweenness and the number of connected components (Table~\ref{tab:gene_solutions}). Three methods yielded more than one connected component: SConES, as described above, SigMod, and HotNet2. HotNet2 produced 135 subnetworks, 115 of which have fewer than five genes. The second largest subnetwork (13 nodes) contained the two breast cancer susceptibility genes \emph{CASP8} and \emph{BLM} (Section~\ref{methods:bcac}).

\begin{table}[!ht]
  \begin{adjustwidth}{-2.25in}{0in} % Comment out/remove adjustwidth environment if table fits in text column.
  \centering
  \caption{
  {\bf Summary statistics on the solutions of multiple network methods on the PPIN. The first row contains the summary statistics on the whole PPIN.}}
  \begin{tabular}{lrrrrrr}
  {\bf Network } & {\bf \# genes } & {\bf \# edges } & {\bf \# components } & {\bf \mean{Betweenness} } & {\bf \median{P}\textsubscript{gene} } & {\bf \# genes in consensus } \\
  \thickhline
  HINT HT           & 13\,619 & 142\,541  & 15  & $1.8 \times 10^{-4}$ & 0.46  & 93/93 \\
  \hline
  dmGWAS            & 194     & 450       & 1   & $5.3 \times 10^{-4}$ & 0.19  & 55/93 \\
  heinz             & 4       & 3         & 1   & $1.2 \times 10^{-3}$ & 0.001 & 4/93  \\
  HotNet2           & 440     & 374       & 130 & $8.3 \times 10^{-5}$ & 0.048 & 63/93 \\
  LEAN              & 0       & 0         & 0   & -         & -     & 0/93  \\
  SConES GI         & 0 (1)   & 0         & 0   & -         & -     & 0/93  \\
  SigMod            & 142     & 249       & 11  & $1 \times 10^{-3}$   & 0.008 & 84/93 \\
  \hline
  Consensus         & 93      & 186       & 21  & $5.5 \times 10^{-4}$ & 0.006 & 93/93 \\
  Stable consensus  & 68      & 49        & 32  & $1 \times 10^{-3}$   & 0.005 & 43/93 \\
  \end{tabular}
  \begin{flushleft} \textbf{\# genes}:~number of genes selected out of those that are part of the PPIN; for SConES GI, the total number of genes, including RNA genes, was added in parentheses. \textbf{\# components}:~number of connected components. \textbf{\mean{Betweenness}}:~median (normalized) betweenness of the selected genes in the PPIN. \textbf{\median{P}\textsubscript{gene}}:~median VEGAS2 P\=/value of the selected genes. \textbf{\# genes in consensus}:~number of genes in common between the method's solution and the 93 genes in the consensus solution.
  \end{flushleft}
  \label{tab:gene_solutions}
  \end{adjustwidth}
\end{table}

@@

Despite their shared properties, the differences between the solutions suggested that each of them captured different aspects of cancer susceptibility. Indeed, out of the 668 genes that were selected by at least one method, only 93 were selected by at least two, 20 by three, and none by four or more. Encouragingly, the more methods selected a gene, the higher its association score to the phenotype (Fig~\ref{fig:solution_overview}B), a relationship that plateaued at 2. Hence, to leverage their strengths and compensate for their respective weaknesses, we built a consensus solution using the genes shared among at least two solutions (Section~\ref{methods:methods}). This solution (Fig~\ref{fig:consensus}A) contained 93 genes and exhibited the aforementioned properties of the individual solutions: enrichment in breast cancer susceptibility genes and higher betweenness centrality than the rest of the genes.

@@

\begin{figure}[!ht]
  \centering
  \includegraphics[width=.8\linewidth]{./figures/figure_4.pdf}
  \caption{\textbf{Drawbacks encountered when using network methods.} \textbf{(A)}~DmGWAS solution, with the genes colored according to the -log$_{10}$ of their P\=/value. \textbf{(B)}~Number of times a gene was selected by either dmGWAS, heinz, LEAN, or SigMod in 100 rewirings of the PPIN (Section~\ref{methods:pathway_enrichment}) and its betweenness. \textbf{(C)}~Betweenness and -log\textsubscript{10} of the VEGAS2 P\=/value in BCAC for each of the nodes in the PPIN. We highlighted the genes selected by each method and the ones selected by more than one (``Consensus''). We labeled the three most central genes that were picked by any method. \textbf{(D)}~Overlap between the solutions of SConES GS, GM, or GI. Barplots are colored based on whether the SNPs map to a gene or not (Section~\ref{methods:snp2gene}).}
  \label{fig:issues}
  \end{figure}

Another peculiarity of network methods was their relationship to betweenness centrality. We studied random rewirings of the PPIN that preserved node centrality (Section~\ref{methods:rewiring}). In this setting, network methods favored central genes (Fig~\ref{fig:issues}B) even though highly central genes often had no association to breast cancer susceptibility (Fig~\ref{fig:issues}C). We found this bias especially in SigMod (\nameref{sfig:rewirings}), which selected three highly central, unassociated genes in both the PPIN and in many of the random rewirings: \emph{COPS5}, \emph{CUL3}, and \emph{FN1}. However, as we showed in Section~\ref{results:consensus} and will show in \ref{results:stable-consensus}, there is evidence in the literature of the contribution of the first two to breast cancer susceptibility. With regards to \emph{FN1}, it encodes a fibronectin, a protein of the extracellular matrix involved in cell adhesion and migration. Overexpression of \emph{FN1} has been observed in breast cancer \cite{Ioachim2002}, and it anticorrelates with poor prognosis in other cancer types \cite{Yi2016,Sponziello2016}.

@@

Most network methods, including the consensus, were highly unstable (Fig~\ref{fig:benchmark}B), raising questions about the results' reliability. We built a new, \emph{stable consensus} solution using the genes selected most often across the 30 solutions obtained by running the 6 methods on 5 different splits of the data (Section~\ref{methods:benchmark}). Such a network should capture the subnetworks more often found altered, and hence should be more resistant to noise. We used only genes selected in at least 7 solutions, which corresponded to 1\% of all genes selected at least once. The resulting stability-based consensus was composed of 68 genes (Fig~\ref{fig:stable-consensus}A). This network shared most of the properties of the consensus: breast cancer susceptibility genes were overrepresented (P\=/value = 3 \texttimes{} 10\textsuperscript{-4}), as well as genes involved in mitochondrial translation and the attenuation phase (adjusted P\=/values 0.001 and 3 \texttimes{} 10\textsuperscript{-5} respectively); the selected genes were more central than average (P\=/value = 1.1 \texttimes{} 10\textsuperscript{-14}); and a considerable number of nodes (19) were isolated (Fig~\ref{fig:stable-consensus}B and C).

@@

\paragraph*{S5 Fig.}
\label{sfig:rewirings}
\textbf{Number of times a gene was selected by either dmGWAS, heinz, or SigMod in 100 rewirings of the PPIN (Section~\ref{methods:rewiring}) and its betweenness.} This figure is equivalent to Fig~\ref{fig:issues}B, split by method. 
%\includegraphics[width=.9\linewidth]{./figures/sfigure_5.pdf

@@

We wish to thank Om Kulkarni for helpful discussion on gene-based GWAS and PPIN databases, Vivien Goepp for his constructive comments of the manuscript, and the genetic epidemiology platform (the PIGE, Plateforme d'Investigation en Génétique et Epidemiologie: S. Eon-Marchais, M. Marcou, D. Le Gal, L. Toulemonde, J. Beauvallet, N. Mebirouk, E. Cavaciuti), the biological resource center (S. Mazoyer, F. Damiola, L. Barjhoux, C. Verny-Pierre, V. Sornin). We wish to pay tribute to Olga M. Sinilnikova, one of the initiators and principal investigators of GENESIS, and who died prematurely on June 30, 2014.
